---
title: " "
output:
   html_document:
     css: style.css
---

<p align = "justify">

<font size="5"> Data analysis for: </font>

<font size="4"> Dylan J. Padilla Perez, 2023. Geographic variation of the *for* gene reveals signatures of local adaptation in *Drosophila melanogaster*.</font>
</p>

</p></p>

[Dylan J. Padilla Perez](https://dylanpadilla.netlify.app/), School of Life Sciences, Arizona State University, Tempe, AZ 85287, USA.
</p></p>


```{r, setup, include=FALSE}

knitr::opts_chunk$set(dpi = 300, warning = FALSE, error = FALSE, message = FALSE, fig.align = "center", fig.height = 7, fig.width = 7)

```


<b>
<font size="5"> Library </font>
</b>

</p></p>

```{r,  echo = TRUE, warning = FALSE, message = FALSE}

library(adegenet)
library(ade4)
library(car)
library(ComplexHeatmap)
library(data.table)
library(dartR)
library(dplyr)
library(ecodist)
library(ggsn)
library(grid)
library(gridBase)
library(ggplot2)
library(gplots)
library(ggpubr)
library(ggspatial)
library(hierfstat)
library(lattice)
library(LEA)
library(lfmm)
library(MASS)
library(maps)
library(maptools)
library(mapplots)
library(miscTools)
library(naniar)
library(OutFLANK)
library(patchwork)
library(pegas)
library(poppr)
library(qvalue)
library(randomcoloR)
library(raster)
library(reshape)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(rworldxtra)
library(rworldmap)
library(scales)
library(sf)
library(shape)
library(SeqArray)
library(SeqVarTools)
library(SNPRelate)
library(stringr)
library(vcfR)
library(vegan)


```

</p></p>
<b>
<font size="5"> Session information </font>
</b>
</p></p>

```{r, warning = FALSE, comment = " "}

R.version
sessionInfo()

```

<br><br>
<b>
<font size="5"> Quality control and SNP isolation </font>
</b>
<br><br>



```{r}


## Setting a seed

set.seed(94)


## load meta-data file

samps <- fread("samps_10Nov2020.csv")

## open GDS for common SNPs (PoolSNP)

genofile <- seqOpen("dest.all.PoolSNP.001.50.10Nov2020.ann.gds", allow.duplicate = TRUE)

## common SNP.dt

snp.dt <- data.table(chr = seqGetData(genofile, "chromosome"),
                     pos = seqGetData(genofile, "position"),
                     nAlleles = seqGetData(genofile, "$num_allele"),
                     id = seqGetData(genofile, "variant.id"),
                     genotype = seqGetData(genofile, "allele"))

snp.dt <- snp.dt[nAlleles == 2]
seqSetFilter(genofile, snp.dt$id)

## filter to target

snp.tmp <- data.table(chr ="2L", pos = 3622086:3656954)
setkey(snp.tmp, chr, pos)
setkey(snp.dt, chr, pos)
seqSetFilter(genofile, variant.id=snp.dt[J(snp.tmp), nomatch = 0]$id)

## get annotations

message("Annotations")
tmp <- seqGetData(genofile, "annotation/info/ANN")
len1 <- tmp$length
len2 <- tmp$data

snp.dt1 <- data.table(len = rep(len1, times = len1), ann = len2, id = rep(snp.dt[J(snp.tmp), nomatch = 0]$id, times = len1))

## Extract data between the 2nd and third | symbol

snp.dt1[ ,class := tstrsplit(snp.dt1$ann,"\\|")[[2]]]
snp.dt1[ ,gene := tstrsplit(snp.dt1$ann,"\\|")[[4]]]

## Collapse additional annotations to original SNP vector length

snp.dt1.an <- snp.dt1[,list(n = length(class), col = paste(class, collapse = ","), gene = paste(gene, collapse = ",")), list(variant.id = id)]

snp.dt1.an[,col := tstrsplit(snp.dt1.an$col,"\\,")[[1]]]
snp.dt1.an[,gene := tstrsplit(snp.dt1.an$gene,"\\,")[[1]]]

## get frequencies

message("Allele Freqs")

ad <- seqGetData(genofile, "annotation/format/AD")
dp <- seqGetData(genofile, "annotation/format/DP")

af <- data.table(ad = expand.grid(ad$data)[,1],
                 dp = expand.grid(dp)[,1],
                 sampleId = rep(seqGetData(genofile, "sample.id"), dim(ad$data)[2]),
                 variant.id = rep(seqGetData(genofile, "variant.id"), each = dim(ad$data)[1]))

## merge them together

message("merge")
afi <- merge(af, snp.dt1.an, by = "variant.id")
afi <- merge(afi, snp.dt, by.x = "variant.id", by.y = "id")

afi[ , af := ad/dp]

## calculate effective read-depth

afis <- merge(afi, samps, by = "sampleId")

afis[chr == "X", nEff := round((dp*nFlies - 1)/(dp+nFlies))]
afis[chr != "X", nEff := round((dp*2*nFlies - 1)/(dp+2*nFlies))]
afis[ ,af_nEff := round(af*nEff)/nEff]


## subsetting dataset

names(afis)
season.dat <- as.data.frame(afis[ , c(21, 30, 1, 13, 14, 15, 19, 18, 20, 2, 11, 12, 6, 22, 99)])
str(season.dat)
head(season.dat)
dim(season.dat)

season.dat$season[season.dat$season == "frost"] <- "fall"

season.dat$season <- as.factor(season.dat$season)
season.filter.dat <- data.frame()

localities <- levels(as.factor(season.dat$locality))
seasons <- c("fall", "spring")

## getting samples collected at least once during spring and winter

for(loc in localities){
    dft <- season.dat[season.dat$locality == loc, ]
    if(all(seasons %in% unique(dft$season))){
        season.filter.dat <- rbind(season.filter.dat, dft)
    }
}

dim(season.filter.dat)
dim(na.omit(season.filter.dat))
str(season.filter.dat)


mat1 <- season.filter.dat[ , c(3, 6, 5, 9, 10, 14, 12, 15)]
names(mat1)
str(mat1)
dim(mat1)

mat1 <- mat1[!is.na(mat1$nEff), ]
dim(mat1)
mat1 <- mat1[mat1$nEff >= 28, ] ## Applying a Neff filter of 28
dim(mat1)
mat1 <- mat1[ , -8]
str(mat1)

## Changing name of cities

mat1$city <- as.factor(mat1$city)
levels(mat1$city)

levels(mat1$city) <- gsub("valday", "Valday", levels(mat1$city))
levels(mat1$city) <- gsub("Odesa", "Odessa", levels(mat1$city))
levels(mat1$city) <- gsub("Charlotttesville", "Charlottesville", levels(mat1$city))
levels(mat1$city) <- gsub("Yesiloz TR13", "Yesiloz", levels(mat1$city))
levels(mat1$city) <- gsub("Yesiloz TR14", "Yesiloz", levels(mat1$city))
levels(mat1$city) <- gsub("Yesiloz TR15", "Yesiloz", levels(mat1$city))
levels(mat1$city) <- gsub("Yesiloz TR16", "Yesiloz", levels(mat1$city))
levels(mat1$city) <- gsub("Yesiloz TR17", "Yesiloz", levels(mat1$city))
levels(mat1$city) <- gsub("Yesiloz TR18", "Yesiloz", levels(mat1$city))
levels(mat1$city) <- gsub("Karensminde Orchard", "Karensminde", levels(mat1$city))
levels(mat1$city) <- gsub("Slankamen. Vinogradi", "Slankamen-Vinogradi", levels(mat1$city))
levels(mat1$city) <- gsub("Chalet \xe0 Gobet", "Chalet-A-Gobet", levels(mat1$city))

levels(mat1$city)
levels(mat1$city)[c(9, 10, 11, 12, 13, 14)] <- "Chornobyl"
levels(mat1$city)
levels(mat1$city)[22] <- "Kyiv"
levels(mat1$city)
levels(mat1$city)[29] <- "Slankamenacki-Vinogradi"
levels(mat1$city)


dim(mat1)
str(mat1)

mat1 <- mat1[!mat1$country == "United Kingdom", ] ## The united kingdom only has 4 pooled samples, so I removed it from the analyses

mat2 <- reshape(mat1, idvar = c("sampleId", "city", "country", "season", "continent"), timevar = "variant.id", direction = "wide")
mat2[1:6, 1:10]
dim(mat2)


mat2$city <- as.character(mat2$city)
mat2$season <- as.character(mat2$season)

mat2 <- mat2[mat2$city != "Homestead", ]
mat2 <- mat2[mat2$city != "Athens", ]
mat2 <- mat2[mat2$city != "Sudbury", ]
mat2 <- mat2[mat2$city != "Brzezina", ]
mat2 <- mat2[mat2$city != "Kalna", ]
mat2 <- mat2[mat2$city != "Slankamenacki-Vinogradi", ]
mat2 <- mat2[mat2$city != "Topeka", ]
mat2 <- mat2[mat2$city != "Chalet-A-Gobet", ]

usa <- c("Lancaster", "Ithaca", "Cross Plains", "Benton Harbor", "Linvilla", "State College", "Tuolumne", "Esparto", "Charlottesville")

for(i in usa){
    mat2$country[mat2$city == i] <- i
}

mat2$country <- as.factor(mat2$country)

levels(mat2$country)
levels(mat2$country)[c(2, 4)] <- "MI-WI"
levels(mat2$country)
levels(mat2$country)[c(9, 10)] <- "NY-MA"
levels(mat2$country)
levels(mat2$country)[c(10, 13)] <- "PA"
levels(mat2$country)


mat3 <- as.data.frame(getGenotypeAlleles(genofile)) ## extracting genotypes from the GDS file
mat3[1:5, 1:5]

for(i in 1:ncol(mat3)){
    mat3[ , i] <- gsub("/", "", mat3[, i])
}

mat3[1:5, 1:5]
mat3$ind <- rownames(mat3)
dim(mat3)
mat3[1:5, 1610:1614]

mat4 <- mat3[mat3$ind %in% mat2$sampleId, ]
dim(mat4)
mat4[1:5, 1600:1614]

same_ord <- mat2[ , c(1, 2, 3)]
colnames(same_ord) <- c("ind", "city", "country")
mat5 <- merge(mat4, same_ord, by = "ind")
dim(mat5)
mat5[1:5, 1:5]
rownames(mat5) <- mat5[ , 1]
dim(mat5)
mat5[1:5, 1610:1616]


mat6 <- mat5[ , -c(1, 1615, 1616)]
dim(mat6)
mat6[1:5, 1:5]
mat6[1:5, 1609:1613]

fly_gen <- df2genind(mat6, ploidy = 2, ind.names = rownames(mat6), pop = mat5$country, sep = "")
fly_gen
fly_gen$tab[1:5, 1:5]
summary(fly_gen$pop)

indmiss_fly <- propTyped(fly_gen, by = "ind")
indmiss_fly[which(indmiss_fly < 0.80)]

barplot(indmiss_fly, ylim = c(0, 1), ylab = "Complete genotypes (proportion)", xlab = "", las = 2, cex.names = 0.4)
mtext("Genotypes", side = 1, line = 3.7)

fly_gen <- missingno(fly_gen, type = "geno", cutoff = 0.20) ## This is the genind file that I will use for all the analyses
fly_gen

## quality control begins here

mlg(fly_gen) ## keep only polymorphic loci
isPoly(fly_gen) %>% summary
poly_loci <- names(which(isPoly(fly_gen) == TRUE))
fly_gen2 <- fly_gen[loc = poly_loci]
isPoly(fly_gen2) %>% summary

fly_gen2$loc.n.all <- fly_gen2$loc.n.all[which(fly_gen2$loc.n.all == 2)]

n <- names(which(nAll(fly_gen2) == 2))
fly_gen2 <- fly_gen2[loc = n]
fly_gen2

indmiss_fly2 <- propTyped(fly_gen2, by = "ind")

barplot(indmiss_fly2, ylim = c(0, 1), ylab = "Complete genotypes (proportion)", xlab = "", las = 2, cex.names = 0.4)
mtext("Genotypes", side = 1, line = 3.7)

locmiss_fly <- propTyped(fly_gen2, by = "loc")
locmiss_fly[which(locmiss_fly < 0.80)] ## print loci with < 80% complete genotypes

barplot(locmiss_fly, ylim = c(0, 1), ylab = "Complete genotypes (proportion)", xlab = "", las = 2, cex.names = 0.4)
mtext("Locus", side = 1, line = 3)

fly_gen3 <- missingno(fly_gen2, type = "loci", cutoff = 0.20)
fly_gen3
locmiss_fly3 <- propTyped(fly_gen3, by = "loc")
locmiss_fly3[which(locmiss_fly3 < 0.80)]
fly_gen3$tab[1:5, 1:5]
min(ploidy(fly_gen3))
max(ploidy(fly_gen3))
summary(fly_gen3$pop)

mlg(fly_gen3)
isPoly(fly_gen3) %>% summary
poly_loci3 <- names(which(isPoly(fly_gen3) == TRUE))
fly_gen3 <- fly_gen3[loc = poly_loci3]
isPoly(fly_gen3) %>% summary

barplot(locmiss_fly3, ylim = c(0, 1), ylab = "Complete genotypes (proportion)", xlab = "", las = 2, cex.names = 0.4)
mtext("Locus", side = 1, line = 3)


## computing basic stats

basic_fly <- basic.stats(fly_gen3, diploid = TRUE)

## testing for H-W equilibrium

## Chi-squared test: p-value

HWE.test <- data.frame(sapply(seppop(fly_gen3), 
                              function(ls) pegas::hw.test(ls, B = 0)[ , 3]))


HWE.test.chisq <- t(data.matrix(HWE.test))
HWE.test.chisq[1:5, 1:5]


## Monte Carlo: p-value

HWE.test <- data.frame(sapply(seppop(fly_gen3), 
                              function(ls) pegas::hw.test(ls, B = 1000)[ , 4]))

HWE.test.MC <- t(data.matrix(HWE.test))
HWE.test.MC[1:5, 1:5]


alpha <- 0.05

Prop.loci.out.of.HWE <- data.frame(Chisq = apply(HWE.test.chisq < alpha, 2, mean), MC = apply(HWE.test.MC < alpha, 2, mean))

Prop.loci.out.of.HWE[1:10, ]


## "False discovery rate" correction for the number of tests. Here I use the function ‘p.adjust’ with the argument method = "fdr" to adjust the p-values from the previous tests

Chisq.fdr <- matrix(p.adjust(HWE.test.chisq, method = "fdr"), 
                    nrow = nrow(HWE.test.chisq))

MC.fdr <- matrix(p.adjust(HWE.test.MC, method = "fdr"), 
                    nrow = nrow(HWE.test.MC))

Prop.loci.out.of.HWE <- data.frame(Chisq = apply(HWE.test.chisq<alpha, 2, mean), 
           MC = apply(HWE.test.MC<alpha, 2, mean),
           Chisq.fdr = apply(Chisq.fdr<alpha, 2, mean),
           MC.fdr = apply(MC.fdr<alpha, 2, mean))

head(Prop.loci.out.of.HWE)


loci <- data.frame()

idx <- 1

for(i in 1:nrow(Prop.loci.out.of.HWE)){
    fdr <- Prop.loci.out.of.HWE[i , ]
    if(fdr$MC.fdr > 0.5){
        loci <- rbind(fdr, loci)
    }

}

loci ## none of the loci are consistenly out of HWE. Some of them are out of HWE in less than 50% of the populations.
dim(loci)


## check for linkage disequilibrium (LD)

LD <- as.genclone(fly_gen3)
LD

quartz()
LD.general <- poppr::ia(LD, sample = 500)
LD.general



```

<br><br>
<b>
<font size="5"> Computing pairwise Fst test </font>
</b>
<br><br>

```{r}

fly_fst <- genet.dist(fly_gen3, method = "WC84") %>% round(digits = 3)
fly_fst[[1:10, 1:10]]

fst.mat <- as.matrix(fly_fst)
fst.mat[fst.mat < 0] <- 0
fst.mat[1:10, 1:10]

tab <- gi2gl(fly_gen3, parallel = FALSE, verbose = NULL) ## converting genind object to genlight object

pval_fst <- gl.fst.pop(tab, nboots = 1000, percent = 95, nclusters = 1, verbose = NULL)

fst <- as.matrix(as.dist(pval_fst$Fsts))
fst[fst < 0] <- 0
fst


```

<br><br>

```{r}

my_palette <- colorRampPalette(c("green", "yellow", "red"))(n = 100)

heatmap.2(fst.mat, density.info = "none", trace = "none", scale = "none", cexRow = 0.7, cexCol = 0.7, key.title = "Fst", srtCol = 45, srtRow = 35, margins = c(8.5, 5), col = my_palette)

```
<p align = "justify">
<font size="4"> **Figure 1.** Heatmap depicting genetic differentiation among populations of *D. melanogaster* based on pairwise Fst values. In some cases, samples from adjacent localities were combined to avoid low sample sizes. Abbreviations are as follows: PA = Pennsylvania; MI-WI = Michigan and Wisconsin; NY-MA = New York and Massachusetts. </font>
</p>
<br><br>


<b>
<font size="5"> OutFlANK outlier analysis </font>
</b>
<br><br>

```{r}

fly_gen3

snps <- as.data.frame(nAll(fly_gen3))
indi <- as.data.frame(fly_gen3$tab)

rownames(snps)[1:10]
rownames(indi)[1:10]

seqSetFilter(genofile, variant.id = rownames(snps), sample.id = rownames(indi))

##seqGDS2VCF(genofile, "my_vcf.gz", verbose = TRUE)

outest <- read.vcfR("my_vcf.gz")

dataout <- vcfR2genind(outest)
dataout$pop <- as.factor(fly_gen3$pop)

outflnk <- gl.outflank(dataout, qthreshold = 0.05, plot = FALSE)

outflnk.df <- outflnk$outflank$results

rowsToRemove <- seq(1, nrow(outflnk.df), by = 2)
outflnk.df <- outflnk.df[-rowsToRemove, ]

outflnk.df$OutlierFlag %>% summary
outflnk.df <- na.omit(outflnk.df)

outflnk.df$FST[outflnk.df$FST < 0] <- 0
head(outflnk.df)

## checking assumptions

plot(outflnk.df$FST, outflnk.df$FSTNoCorr, pch = 19, col = alpha("black", 0.3), ylab = "Fst uncorrected", xlab = "Fst corrected", las = 1, main = "Loci deviation from the linear relationship")
abline(0, 1, col = "skyblue", lwd = 2)


```

<br><br>

```{r}


## Italic labels

fstlab <- expression(italic("F")[ST])
hetlab <- expression(italic("H")[e])


## Plot He versus Fst

outFlank <- ggplot(data = outflnk.df) +
  geom_point(aes(x = He, y = FST, colour = OutlierFlag)) +
  scale_colour_manual(values = c("black","red"), labels = c("Neutral SNP","Outlier SNP")) +
  xlab(hetlab) +
  ylab(fstlab) +
  theme(legend.title = element_blank(),
  )


outFlank

```
<p align = "justify">
<font size="4"> **Figure 3.** Manhattan plot of the distribution of Fst loci as a function of their heterozygosity. No Fst loci showed significant deviation from the genetic background. </font>
</p>
<br><br>


<b>
<font size="5"> Genetic admixture proportions </font>
</b>
<br><br>


```{r}


## exploring the data once again (do it all the time!)

fly_gen3
fly_gen3$tab[1:5, 1:5]
nLoc(fly_gen3) # number of loci
nPop(fly_gen3) # number of sites
nInd(fly_gen3) # number of individuals
summary(fly_gen3$pop) # sample size
min(ploidy(fly_gen3))
max(ploidy(fly_gen3))


snmf1 <- load.snmfProject("gl_geno.snmfProject")


## extract Q-matrix for the best run

plot(snmf1, col = "blue", cex = 1.5, pch = 19, las = 1)


## extract the cross-entropy of all runs where K = 9

ce <- cross.entropy(snmf1, K = 9)
ce


## find the run with the lowest cross-entropy

lowest.ce <- which.min(ce)
lowest.ce


## extract Q-matrix for the best run

qmatrix <- as.data.frame(Q(snmf1, K = 9, run = lowest.ce))
head(qmatrix)

## changing order of levels

pops <- fly_gen3$pop
levels(pops)
levels(pops)[3] <- "Charlott."


qmplot <- cbind(qmatrix, pops)
qmplot$pops <- factor(qmplot$pops, levels = c("Esparto", "Tuolumne", "MI-WI", "Charlott.", "PA", "NY-MA", "Spain", "France", "Germany", "Austria", "Denmark", "Finland", "Ukraine", "Russia", "Turkey"))
qmplot <- qmplot[order(qmplot$pops), ]

pops <- qmplot$pops

```




```{r}

layout(matrix(c(0, 1, 1, 0,
                0, 1, 1, 0,
                2, 2, 2, 2,
                2, 2, 2, 2), nrow = 4, ncol = 4, byrow = TRUE))

par(mar = c(4, 4, 1, 0))
plot(snmf1, col = "blue", cex = 1.5, pch = 19, las = 1, cex.lab = 1.3)
Arrows(x = 9, y = 0.295, x1 = 9, y1 = 0.305, col = "red", arr.type = "triangle", code = 1, lwd = 1.5, arr.length = 0.2)
mtext("(a)", side = 2, at = 0.345, line = 2.8, font = 2, family = "serif", las = 1)

par(mar = c(5, 4.5, 1, 0))
barplot(t(qmplot[1:9]), col = RColorBrewer::brewer.pal(9,"Paired"), border = NA, space = 0, xlab = "", xaxt = "n",  ylab = "Admixture proportion", las = 1, cex.lab = 1.4)

## adding population labels to the axis:

names <- unique(as.character(pops))

medians <- c()

for(i in 1:length(pops)){
    
    axis(1, at = median(which(pops == pops[i])), labels = "")
    medians <- c(medians, median(which(pops == pops[i])))
}

medians[1:10]

text(x = as.numeric(unique(as.character(medians))), y = par("usr")[3] - 0.05, labels = names, xpd = NA, srt = 35, cex = 1, adj = 1.2)
mtext("Individuals", side = 1, line = 4)
mtext("(b)", side = 2, at = 1.1, line = 2.4, font = 2, family = "serif", las = 1)

```

<p align = "justify">
<font size="4"> **Figure 4.** Population structure analysis based on individual ancestry coefficients for a number of ancestral populations. **(a)** Cross-entropy values for each *snmf* run with *k* ranging between *k=1* and *k=10*. The red arrow indicates the most likely value of *k*. **(b)** Admixture proportion across populations of *D. melanogaster*. Colors indicate genetic clusters. </font>
</p>
<br><br>


```{r}

## label column names of qmatrix

ncol(qmatrix)
cluster_names = c()

for(i in 1:ncol(qmatrix)){
  cluster_names[i] = paste("Cluster", i)
}

cluster_names
colnames(qmatrix) <- cluster_names
head(qmatrix)
dim(qmatrix)


## add individual IDs

qmatrix$Ind <- indNames(fly_gen3)

## add site IDs

qmatrix$Site <- fly_gen3$pop
head(qmatrix)

## calculate mean admixture proportions for each site

clusters <- grep("Cluster", names(qmatrix)) ## indexes of cluster columns
avg_admix <- aggregate(qmatrix[, clusters], list(qmatrix$Site, qmatrix$Ind), mean)
colnames(avg_admix)[1:2] <- c("country", "Ind")
head(avg_admix)
str(avg_admix)



## import csv file containing coordinates

coor <- read.csv("coor.csv")
str(coor)
head(coor)


admix <- merge(coor, avg_admix, by = "country")
str(admix)
head(admix)
colnames(admix)[c(5, 6, 7, 8, 9, 10, 11, 12, 13)] <- c("cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "cluster6", "cluster7", "cluster8", "cluster9")


```


<br><br>
<b>
<font size="5"> Plotting admixture proportion on a map </font>
</b>
<br><br>

```{r}


## Plotting map

quartz()

map("world", fill = TRUE, col = "gray", xlim = c(-130, 50), ylim = c(-20, 70), border = 0, mar = c(0, 5.5, 0, 0))

map.axes(cex.axis = 0.7, las = 1)
mtext(side = 1, line = 2, "Longitude")
mtext(side = 2, line = 2, "Latitude")


## Adding pie charts to the map

for(i in 1:nrow(admix)){
    add.pie(z = c(admix$cluster1[i], admix$cluster2[i], admix$cluster3[i], admix$cluster4[i], admix$cluster5[i], admix$cluster6[i], admix$cluster7[i], admix$cluster8[i], admix$cluster9[i]), x = admix$long[i], y = admix$lat[i], radius = 2, col = RColorBrewer::brewer.pal(9,"Paired"), labels = "", border = TRUE)
    
}

```
<p align = "justify">
<font size="4"> **Figure 5.** Mean admixture proportions across populations of *D. melanogaster* surveyed in America and Europe. Colors in the pie charts represent genetic clusters. </font>
</p>
<br><br>




<br><br>
<b>
<font size="5"> Isolation by Environment and Isolation by Distance analyses </font>
</b>
<br><br>


```{r}

unique(fly_gen3$pop)

obj <- matrix(NA, nrow = length(unique(fly_gen3$pop)), ncol = 3, byrow = TRUE)

idx <- 1

for(i in unique(fly_gen3$pop)){
    long <- admix$long[admix$country == i][1]
    lat <- admix$lat[admix$country == i][1]
    country <- i
    obj[idx, ] <- c(country, long, lat)
    idx = idx + 1

}

obj

obj <- as.data.frame(obj)
colnames(obj) <- c("country", "x", "y")
obj$x <- as.numeric(obj$x)
obj$y <- as.numeric(obj$y)
rownames(obj) <- obj[ , 1]
obj <- obj[ , c(2, 3)]
str(obj)

fly_gen3@other <- obj[ , c(1, 2)]
fly_gen3

## genetic distance

GD.pop.PairwiseFst.hierfstat <- as.dist(hierfstat::pairwise.neifst(hierfstat::genind2hierfstat(fly_gen3)))

## geographic distance

Dgeo <- round(dist(fly_gen3$other), 3)

## making sure the order of rows and columns are the same in all of the matrices

ord.gen <- as.matrix(GD.pop.PairwiseFst.hierfstat)
ord.geo <- as.matrix(Dgeo)

ord.geo
ord.gen <- ord.gen[rownames(ord.geo), colnames(ord.geo)]
ord.gen

fst.dist <- round(as.dist(ord.gen), 3)
fst.dist
Dgeo


NPP_file <- list.files("/Users/dylanpadilla/Dropbox (ASU)/npp-geotiff", full.names = TRUE)
NPP_file

NPP.data <- stack(NPP_file)

NPP.data.ext.dis <- extract(NPP.data, obj)
head(NPP.data.ext.dis)

rownames(NPP.data.ext.dis) <- rownames(obj)

DNPP <- round(dist(log(NPP.data.ext.dis)), 3)
DNPP


## partial mantel test

part.mant <- mantel.partial(fst.dist, DNPP, Dgeo, method = "pearson", permutations = 1000)
part.mant

part.mant2 <- mantel.partial(fst.dist, Dgeo, DNPP, method = "pearson", permutations = 1000)
part.mant2

## mantel test

ibd <- mantel.randtest(fst.dist, Dgeo)
ibd

quartz()
plot(ibd, las = 1, main = "") ## isolation by distance is clearly significant

mantCor <- mgram(fst.dist, Dgeo, nperm = 1000)
mantCor

mantCor2 <- mgram(fst.dist, DNPP, nperm = 1000)
mantCor2

plot(mantCor, las = 1)
plot(mantCor2, las = 1)


## filled dots indicate significant correlations. The Mantel correlogram shows that genotypes are relatively similar at short distance, while this similarity decreases with distance. The negative correlations indicate genetic differenciation between some populations


dens <- MASS::kde2d(as.vector(Dgeo), as.vector(fst.dist), n = 300)
myPal <- colorRampPalette(c("white", "blue", "gold", "orange", "red"))
plot(as.vector(Dgeo), as.vector(fst.dist), pch = 20, xlab = "Geographic Distance", ylab = "Genetic Distance", las = 1)
image(dens, col = transp(myPal(300), 0.7), add = TRUE)
abline(lm(as.vector(fst.dist) ~ as.vector(Dgeo)), lwd = 2)
lines(loess.smooth(as.vector(Dgeo), as.vector(fst.dist)), col = "red", lwd = 2)


dens2 <- MASS::kde2d(as.vector(DNPP), as.vector(fst.dist), n = 300)
myPal2 <- colorRampPalette(c("white", "blue", "gold", "orange", "red"))
plot(as.vector(DNPP), as.vector(fst.dist), pch = 20, xlab = "NPP Distance", ylab = "Genetic Distance", las = 1)
image(dens2, col = transp(myPal2(300), 0.7), add = TRUE)
abline(lm(as.vector(fst.dist) ~ as.vector(DNPP)), lwd = 2)
lines(loess.smooth(as.vector(DNPP), as.vector(fst.dist)), col = "red", lwd = 2)



```


```{r}

layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                3, 3, 4, 4,
                3, 3, 4, 4), nrow = 4, ncol = 4, byrow = TRUE))


par(mar = c(4, 4.1, 1, 1))
plot(as.vector(Dgeo), as.vector(fst.dist), pch = 20, xlab = "Geographic Distance", ylab = "Genetic Distance", las = 1, cex.lab = 1.4)
image(dens, col = transp(myPal(300), 0.7), add = TRUE)
abline(lm(as.vector(fst.dist) ~ as.vector(Dgeo)), lwd = 2)
lines(loess.smooth(as.vector(Dgeo), as.vector(fst.dist)), col = "red", lwd = 2)
mtext("(a)", side = 2, at = 0.07, line = 2.6, font = 2, family = "serif", las = 1)

par(mar = c(4, 4.1, 1, 1))
plot(as.vector(DNPP), as.vector(fst.dist), pch = 20, xlab = "NPP Distance", ylab = "Genetic Distance", las = 1, cex.lab = 1.4)
image(dens2, col = transp(myPal2(300), 0.7), add = TRUE)
abline(lm(as.vector(fst.dist) ~ as.vector(DNPP)), lwd = 2)
lines(loess.smooth(as.vector(DNPP), as.vector(fst.dist)), col = "red", lwd = 2)
mtext("(b)", side = 2, at = 0.07, line = 3.1, font = 2, family = "serif", las = 1)

par(mar = c(4, 4.1, 1, 1))
plot(mantCor, las = 1, cex.axis = 1, cex.lab = 1.6)
mtext("(c)", side = 2, at = 0.7, line = 2.6, font = 2, family = "serif", las = 1)

par(mar = c(4, 4.1, 1, 1))
plot(mantCor2, las = 1, cex.axis = 1, cex.lab = 1.6, ylim = c(-0.4, 0.6))
mtext("(d)", side = 2, at = 0.7, line = 3.1, font = 2, family = "serif", las = 1)


```
<p align = "justify">
<font size="4"> **Figure 2.** Patterns of isolation by distance and isolation by environment according to Partial Mantel tests and Correlograms. **(a-b)** Geographic and net primary production distances as a function of genetic distance. Colors represent estimated probability densities and the red line a smoothed local mean. **(c-d)** Mantel correlation at different distance classes as a function of geographic and net primary production distances. Filled dots indicate significant correlations. </font>
</p>
<br><br>



<b>
<font size="5"> Detecting adaptation </font>
</b>
<br><br>


```{r}

## extracting environmental variables from WorldClim and other sources


mat_adap <- admix[ , c(1, 2, 3, 4)]
head(mat_adap)

## extracting NPP


NPP_file <- list.files("/Users/dylanpadilla/Dropbox (ASU)/npp-geotiff", full.names = TRUE)
NPP_file

NPP.data <- stack(NPP_file)

prod <- mat_adap[ , c(2, 3)] ## loading the dataframe with the coordinates
head(prod)

NPP.data.ext <- extract(NPP.data, prod)
head(NPP.data.ext)


NPP.data.ext <- as.data.frame(NPP.data.ext)
colnames(NPP.data.ext) <- "NPP"
head(NPP.data.ext)
##is.na(NPP.data.ext)


## extracting BIO15 precipitation seasonality (coefficient of variation), BIO4 temperature seasonality (standard deviation * 100)

files <- list.files("/Users/dylanpadilla/Dropbox (ASU)/Padilla et al/wc2.1_30s_bio", full.names = TRUE)
files

pre_sea <- stack(files[4])
tem_sea <- stack(files[10])

coord <- mat_adap[ , c(2, 3)]
head(coord)

prec.season <- extract(pre_sea, coord)
temp.season <- extract(tem_sea, coord)

variables <- cbind(prec.season, temp.season, log(NPP.data.ext))
head(variables)
colnames(variables) <- c("precseason", "tempseason", "logNPP")
head(variables)

env <- cbind(mat_adap, variables)
head(env)
str(env)



## confirm that genotypes and environmental data are in the same order

x <- tab(fly_gen3, NA.method = "mean")

snppca <- as.data.frame(x)
snppca$Ind <- rownames(x)
snppca$country <- fly_gen3$pop
dim(snppca)

gen <- snppca[order(snppca$country), ]

merging <- merge(gen, env, by = "Ind")
merging[1:5, 1:5]
rownames(merging) <- merging[ , 1]
dim(merging)
merging[1:5, 3101:3108]
merging <- merging[ , -c(1, 3102, 3103, 3104, 3105, 3106, 3107, 3108)]
merging[1:5, 3085:3100]
dim(merging)

genomic <- merging

env$Ind <- as.factor(env$Ind)
env$Ind <- as.character(levels(env$Ind))
dim(snppca)
snppca[1:5, 1:5]
snppca[1:5, 3085:3102]

snppca <- snppca[ , -c(3101, 3102)]
snppca[1:5, 3085:3100]
dim(snppca)

identical(rownames(genomic), env[ , 4]) ## genotypes and environmental data are in the same order

names(env)

env.pca <- rda(env[ , c(2, 3, 5, 6, 7)], scale = T)
summary(env.pca)$cont

quartz()
par(las = 1)
screeplot(env.pca, main = " ")

round(scores(env.pca, choices = 1:5, display = "species", scaling = 0), digits = 3)

pred.PC1 <- scores(env.pca, choices = 1, display = "sites", scaling = 0)


## determine K

## I’ll use a broken stick criterion to determine K

par(las = 1)
screeplot(env.pca, bstick = TRUE, type = "lines", las = 1, main = "")

## genomic data

gen.pca <- rda(snppca, scale = T)

par(las = 1)
screeplot(gen.pca, main = " ", bstick = TRUE, type = "lines")

## for the genomic data, I can see that two of the PCs have eigenvalues greater than random (greater than the broken stick values in red). This effectively means that K = 2 for the dataset, based on a PCA assessment

K <- 2:9


fly.lfmm1 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[1]) ## c.ange K as you see fit
fly.lfmm2 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[2]) ## c.ange K as you see fit
fly.lfmm3 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[3]) ## c.ange K as you see fit
fly.lfmm4 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[4]) ## c.ange K as you see fit
fly.lfmm5 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[5]) ## c.ange K as you see fit
fly.lfmm6 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[6]) ## c.ange K as you see fit
fly.lfmm7 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[7]) ## c.ange K as you see fit
fly.lfmm8 <- lfmm_ridge(Y = snppca, X = pred.PC1, K = K[8]) ## c.ange K as you see fit

fly.pv1 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm1, calibrate = "gif")
fly.pv2 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm2, calibrate = "gif")
fly.pv3 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm3, calibrate = "gif")
fly.pv4 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm4, calibrate = "gif")
fly.pv5 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm5, calibrate = "gif")
fly.pv6 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm6, calibrate = "gif")
fly.pv7 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm7, calibrate = "gif")
fly.pv8 <- lfmm_test(Y = snppca, X = pred.PC1, lfmm = fly.lfmm8, calibrate = "gif")

names(fly.pv1) # this object includes raw z-scores and p-values, as well as GIF-calibrated scores and p-values


## let’s look at the genomic inflation factor (GIF)

fly.pv1$gif
fly.pv2$gif
fly.pv3$gif
fly.pv4$gif
fly.pv5$gif
fly.pv6$gif
fly.pv7$gif
fly.pv8$gif

## an appropriately calibrated set of tests will have a GIF of around 1

## let’s look at how application of the GIF to the p-values impacts the p-value distribution


quartz()
layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                3, 3, 4, 4,
                3, 3, 4, 4,
                5, 5, 6, 6,
                5, 5, 6, 6,
                7, 7, 8, 8,
                7, 7, 8, 8), nrow = 8, ncol = 4, byrow = TRUE))

par(mar = c(5, 5, 0.2, 1), las = 1)

hist(fly.pv1$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 2", bty = "n")
hist(fly.pv1$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 2", bty = "n")
hist(fly.pv2$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 3", bty = "n")
hist(fly.pv2$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 3", bty = "n")
hist(fly.pv3$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 4", bty = "n")
hist(fly.pv3$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 4", bty = "n")
hist(fly.pv4$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 5", bty = "n")
hist(fly.pv4$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 5", bty = "n")


quartz()
layout(matrix(c(1, 1, 2, 2,
                1, 1, 2, 2,
                3, 3, 4, 4,
                3, 3, 4, 4,
                5, 5, 6, 6,
                5, 5, 6, 6,
                7, 7, 8, 8,
                7, 7, 8, 8), nrow = 8, ncol = 4, byrow = TRUE))

par(mar = c(5, 5, 0.2, 1), las = 1)

hist(fly.pv5$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 6", bty = "n")
hist(fly.pv5$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 6", bty = "n")
hist(fly.pv6$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 7", bty = "n")
hist(fly.pv6$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 7", bty = "n")
hist(fly.pv7$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 8", bty = "n")
hist(fly.pv7$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 8", bty = "n")
hist(fly.pv8$pvalue[ , 1], xlab = "Unadjusted p-values", las = 1, main = "", ylim = c(0, 500))
legend("top", "K = 9", bty = "n")
hist(fly.pv8$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", "K = 9", bty = "n")



## convert the adjusted p-values to q-values. q-values provide a measure of each SNP’s significance, automatically taking into account the fact that thousands are simultaneously being tested. I can then use an FDR threshold to control the number of false positive detections (given that the p-value distribution is “well-behaved”)

fly.qv1 <- qvalue(fly.pv1$calibrated.pvalue)$qvalues
length(which(fly.qv1 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv2 <- qvalue(fly.pv2$calibrated.pvalue)$qvalues
length(which(fly.qv2 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv3 <- qvalue(fly.pv3$calibrated.pvalue)$qvalues
length(which(fly.qv3 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv4 <- qvalue(fly.pv4$calibrated.pvalue)$qvalues
length(which(fly.qv4 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv5 <- qvalue(fly.pv5$calibrated.pvalue)$qvalues
length(which(fly.qv5 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv6 <- qvalue(fly.pv6$calibrated.pvalue)$qvalues
length(which(fly.qv5 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv7 <- qvalue(fly.pv7$calibrated.pvalue)$qvalues
length(which(fly.qv5 < 0.05)) ## how many SNPs have an FDR < 5%?

fly.qv8 <- qvalue(fly.pv8$calibrated.pvalue)$qvalues
length(which(fly.qv5 < 0.05)) ## how many SNPs have an FDR < 5%?




## using K from 2 to 9, the default GIF correction, and an FDR threshold of 0.05, I detected 2 candidate SNPs under selection in response to the PC1 environmental predictor

(fly.FDR.5 <- colnames(snppca)[which(fly.qv5 < 0.05)]) ## identify which SNPs these are


## visualizing results


## Manhattan plot with causal loci
 
qvalues <- fly.qv5

row <- as.data.frame(qvalues)
str(row)

rownames(row) <- 1:3100

(causal.set <- as.numeric(rownames(row)[which(row < 0.05)]))
(fly.FDR.5 <- gsub("[^0-9]", "", fly.FDR.5))


quartz()
plot(-log10(qvalues), pch = 19, cex = 1, bg = "black", col = alpha("gray", 0.3), xlab = "SNP", las = 1)
points(causal.set, -log10(qvalues)[causal.set], pch = 19, col = alpha("red", 0.3))
text(causal.set, (-log10(qvalues)[causal.set] - 0.05), fly.FDR.5, col = "red")



```


```{r}


quartz()
layout(matrix(c(0, 1, 1, 0,
                0, 1, 1, 0,
                0, 2, 2, 0,
                0, 2, 2, 0), nrow = 4, ncol = 4, byrow = TRUE))

par(las = 1, mar = c(4, 4, 1, 0.6))
hist(fly.pv5$calibrated.pvalue[ , 1], xlab = "GIF-adjusted p-values", las = 1, main = "", ylim = c(0, 350))
legend("top", paste("K = 6\nGIF = ", round(fly.pv5$gif, 2)), bty = "n")
mtext("(a)", side = 2, at = 370, line = 2.5, las = 1, font = 2, family = "serif")


plot(-log10(qvalues), pch = 19, cex = 1, bg = "gray", col = alpha("black", 0.1), xlab = "SNP", las = 1)
points(causal.set, -log10(qvalues)[causal.set], pch = 19, col = alpha("red", 0.3))
text(causal.set, (-log10(qvalues)[causal.set] - 0.05), fly.FDR.5, col = "red")
mtext("(b)", side = 2, at = 1.45, line = 2.5, las = 1, font = 2, family = "serif")



```
<p align = "justify">
<font size="4"> **Figure 6.** Genotype-environment association test based on a latent factor mixed model with ridge penalty. **(a)** Distribution of adjusted p-values using the default genomic inflation factor. **(b)** Manhattan plot of loci (SNPs) potentially affected by the PC1 predictor variable. Loci highlighted in red were considered to be under selection according to a False Discovery Rate of 0.05. </font>
</p>
<br><br>
